MCC Kids Model Predictions

### 2019-03-18 papers/full_paper/models.Rmd
update noisy models

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(coda)
library(ggpubr)

@@

```{r model rsa noise ex3 adults}
predictions.rsa.ex3 <-'

@@

@@

@@

ex_3_pred <- bind_rows(
  ex_3_model_pred,
  adult_ex3_data_summary
)

@@

```{r log likes ex3 adults}
# calculating model log-likelihoods
log_like_pragmatic_ex3 <- adult_ex3_data %>%
  left_join(ex3_pragm_model_predictions)%>%
  mutate(like = ifelse(correct_inf == 1, value, 1 - value),
         log_like = log(like))%>%
  group_by(model)%>%
  summarise(model_log_like = sum(log_like))

@@

>>>>

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(coda)
library(matrixStats)
library(ggpubr)

@@

#### Pragmatic model

##### Parameter free

```{r model rsa ex3 adults}
predictions.rsa.ex3 <-'

@@

##### Noise parameter

```{r model rsa noise ex3, include = FALSE}
predictions.rsa.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

 map(function(d){ 
      observe(noisyModelPredictions, d.correct_inf)
    }, conditionSpecificData)

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

}, allParams)

  return extend(_.fromPairs(conditionOutput), {noise: noise})
}
'
```

```{r predictions pragmatic model ex 3, include = FALSE}
ex3_pragm_model_noise <- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex3.noise, sep='\n'),
  data = list(allData = adult_ex3_data, allParams = model_params_ex3),
  data_var = "dataFromR",
  model_var = "model",
  chains = 4,
  cores = 4,
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
) %>%
  saveRDS("../../stats/saves/ex3_pragm_model_noise.rds")

# model predictions
ex3_pragm_model_pred_noise <- readRDS("../../stats/saves/ex3_pragm_model_noise.rds") %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation","speaker", "alignment"), sep="/")%>%
  mutate(model="pragmatic_noise")%>%
  group_by(model,common_ground_manipulation,speaker,alignment)%>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

## noise parameter
ex3_pragm_model_noise <- readRDS("../../stats/saves/ex3_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))
```
#### Prior only model

@@

#### Flat prior model

@@

ex_3_pred <- bind_rows(
  ex_3_model_pred,
  ex3_pragm_model_pred_noise,
  adult_ex3_data_summary
)

@@

### Loglike Parameter free model
```{r log likes ex3 adults}
# calculating model log-likelihoods
log_like_pragmatic_ex3 <- adult_ex3_data %>%
  left_join(ex3_pragm_model_predictions)%>%
  mutate(like = ifelse(correct_inf == 1, value, 1 - value),
         log_like = log(like))%>%
  group_by(model)%>%
  summarise(model_log_like = sum(log_like))

@@

### Loglike models with noise

```{r model rsa noise ex3, include = FALSE}
loglike.rsa.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

   var noise = uniformDrift({a: 0, b:1, width: 0.1})

  var conditionOutput = map(function(conditionInfo){

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors = [1 - conditionInfo.prior, 1 - conditionInfo.prior, conditionInfo.prior]

      var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors, conditionInfo.speak_opt)

   var noisyModelPredictions = addNoise(modelPredictions, noise)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )
  return  loglike

}, allParams)

 return sum(conditionOutput)

```{r}
# ex3_pragm_model_noise_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.rsa.ex3.noise, sep='\n'),
#   data = list(allData = adult_ex3_data, allParams = model_params_ex3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 2)
# ) %>%
#   saveRDS("../../stats/saves/ex3_pragm_model_noise_loglike.rds")

ex3_pragm_model_noise_loglike <- readRDS("../../stats/saves/ex3_pragm_model_noise_loglike.rds")%>%
  mutate(model = "pragmatic")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```

### 2019-03-18 stats/kids/kids_model_predictions.Rmd
update noisy models

ex3_data_param <- ex3_data%>%
  mutate(different_speaker_pref_int = model_params$different_speaker_pref_int,
         different_speaker_pref_slope = model_params$different_speaker_pref_slope,
         so_int = model_params$NA_so_int,
         so_slope = model_params$NA_so_slope,
         so_slope_quad = model_params$NA_so_slope_quad,
         same_speaker_pref_int = model_params$same_speaker_pref_int,
         same_speaker_pref_slope = model_params$same_speaker_pref_slope,
         so = so_int+so_slope*(age_num-min(age_num))+so_slope_quad*(age_num-min(age_num))^2,
         diff_prior = plogis(different_speaker_pref_int+different_speaker_pref_slope*(age_num-min(age_num))),
         same_prior = plogis(same_speaker_pref_int+same_speaker_pref_slope*(age_num-min(age_num))))%>%
  group_by(subid,trial)%>%
        mutate(priorProbs = ifelse(Speaker == "same_speaker" & Alignment == "congruent", list(c(1-same_prior,1-same_prior,same_prior)),
                                   ifelse(Speaker == "different_speaker" & Alignment == "congruent", list(c(1-diff_prior,1-diff_prior,diff_prior)),
                                          ifelse(Speaker == "same_speaker" & Alignment == "incongruent", list(c(same_prior,same_prior,1-same_prior)),
                                                 list(c(diff_prior,diff_prior,1-diff_prior))

@@

    return  [row.age_num + "/" + row.Speaker + "/" + row.Alignment, Math.exp(noisyModelPredictions.score(1))]

@@

noisy_pred <- noisy_model_pred %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("age", "Speaker", "Alignment"), sep="/")%>%
  mutate( model = "noise parameter across age")%>%
  group_by(model, age,Alignment,Speaker) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))

@@

    return  [row.age_num + "/" + row.Speaker + "/" + row.Alignment, Math.exp(noisyModelPredictions.score(1))]

@@

```{r}
dev_noisy_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, developmentalNoisyModelPred , sep='\n'),
  data =list(ex3data = ex3_data_param),
  data_var = "dataFromR",
  model_var = "model",
  chains = 4,
  cores = 4,
  inference_opts = list(method = "MCMC", samples = 8000, burn = 2000, verbose = T)
)

@@

models <- bind_rows(
  quad_pred,
  lin_pred,
  lin_flat_pred,
  lin_pred_prior
)

p_data <- ex3_data %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  mutate(age_bin = age_num - 3)%>%
  group_by(Speaker, Alignment ,age_bin, subid)

ggplot(models)+
  #geom_jitter(data = p_data, aes(x = age_bin, y = correct_inf), width = .025,height = .025, alpha = .2)+
  #geom_ribbon(aes(x = age, ymin = ci_lower, ymax = ci_upper, fill = model), alpha = .4) +
  geom_line(aes(x = age, y = mean, col = model), size = 1, alpha = 1, linetype = 1) + 
  geom_smooth(data = p_data, aes(x = age_num, y = correct_inf), col = "black", method = "loess", span = 2, se = T,  alpha = .5, size = .4)+
  #geom_hline(yintercept = 0.5, lty=2)+
  labs(y="proportion more informative")+
  ylim(-0.05,1.05)+
  facet_grid(Speaker~Alignment)+
  theme_few()+
  scale_color_solarized()+
  scale_fill_solarized()

@@

>>>>

model_params <- readRDS("../../stats/saves/child_model_param.rds")%>%
  select(value) %>%
  map_df(bind_rows)%>%
  group_by(param,condition)%>%
  summarise(param_map = estimate_mode(val))%>%
  mutate(cond = paste(param,condition, sep = "_"))%>%
  ungroup()%>%
  select(cond,param_map)%>%
  spread(cond,param_map)%>%
  rename(so_int = so_int_NA,
         so_slope = so_slope_NA)

ex3_data_param <- ex3_data%>%
  mutate(different_speaker_pref_int = model_params$pref_int_different_speaker,
         different_speaker_pref_slope = model_params$pref_slope_different_speaker,
         so_int = model_params$so_int,
         so_slope = model_params$so_slope,
         same_speaker_pref_int = model_params$pref_int_same_speaker,
         same_speaker_pref_slope = model_params$pref_slope_same_speaker,
         so = so_int+so_slope*(age_num-min(age_num)),
         diff_prior = plogis(different_speaker_pref_int+different_speaker_pref_slope*(age_num-min(age_num))),
         same_prior = plogis(same_speaker_pref_int+same_speaker_pref_slope*(age_num-min(age_num))))%>%
  group_by(subid,trial)%>%
        mutate(priorProbs = ifelse(Speaker == "same_speaker" & Alignment == "congruent", list(c(1-same_prior,1-same_prior,same_prior)),
                                   ifelse(Speaker == "different_speaker" & Alignment == "congruent", list(c(1-diff_prior,1-diff_prior,diff_prior)),
                                          ifelse(Speaker == "same_speaker" & Alignment == "incongruent", list(c(same_prior,same_prior,1-same_prior)),
                                                 list(c(diff_prior,diff_prior,1-diff_prior))

@@

    return  [row.age_num + "/" + row.Speaker + "/" + row.Alignment + "/" + row.correct_inf, Math.exp(noisyModelPredictions.score(1))]

@@

noisy_pred <- noisy_model_pred %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("age", "speaker", "alignment", "correct"), sep="/")%>%
  mutate( model = "noise parameter across age")%>%
  group_by(model, age,alignment,speaker, correct) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))

log_like_noisy_model <- noisy_pred %>%
  mutate(like = ifelse(correct == "1",mean,1-mean),
         log_like = log(like))%>%
  group_by(model)%>%
  summarise(model_log_like = sum(log_like))

@@

    return  [row.age_num + "/" + row.Speaker + "/" + row.Alignment , Math.exp(noisyModelPredictions.score(1))]

@@

```{r}
dev_noisy_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, developmentalNoisyModelPred , sep='\n'),
  data =list(ex3data = ex3_data_param),
  data_var = "dataFromR",
  model_var = "model",
  chains = 4,
  cores = 4,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 2000, verbose = T)
)

@@

child_ex3_pragm_model_predictions<- child_ex3_pragm_model_predictions%>%
  mutate(mean = value,
         ci_lower = value,
         ci_upper = value)

models <- bind_rows(
  noisy_pred,
child_ex3_pragm_model_predictions
)

p_data <- child_ex3_data %>%
  mutate(age_bin = age_num - 3)%>%
  group_by(speaker, alignment ,age_bin, id)

ggplot(models)+
  #geom_jitter(data = p_data, aes(x = age_bin, y = correct_inf), width = .025,height = .025, alpha = .2)+
  #geom_ribbon(aes(x = age, ymin = ci_lower, ymax = ci_upper, fill = model), alpha = .4) +
  geom_line(aes(x = age, y = mean, col = model), size = 1, alpha = 1, linetype = 1) + 
  geom_smooth(data = p_data, aes(x = age_num, y = correct_inf), col = "black", method = "loess", span = 2, se = T,  alpha = .5, size = .4)+
  #geom_hline(yintercept = 0.5, lty=2)+
  labs(y="proportion more informative")+
  ylim(-0.05,1.05)+
  facet_grid(speaker~alignment)+
  theme_few()+
  scale_color_solarized()+
  scale_fill_solarized()

@@

# model comparison

```{r}
noisyModelPredWebpplLoglike <-'

    display(JSON.stringify(row.correct_inf))

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n:row.correct_inf})

  return  loglike

  return sum(output)

noisy_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, noisyModelPredWebpplLoglike , sep='\n'),
  data =list(ex3data = ex3_data_param),
  data_var = "dataFromR",
  model_var = "model",
  chains = 4,
  cores = 4,
  inference_opts = list(method = "forward", samples = 2)
)

# ex4_pragm_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.rsa.ex4, sep='\n'),
#   data = list(Data = adult_ex4_data, allConditions = allConditionsEx4, predPrior = predPriorEx4, noise = ex3_pragm_model_noise$mean),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex4_pragm_model_loglike.rds")

ex4_pragm_model_loglike <- readRDS("../../stats/saves/ex4_pragm_model_loglike.rds")%>%
  mutate(model = "pragmatic")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))
```
